---
title: "Covid91 vaccine study Final2021"
author: "Thomas Matthews!"
date: "May 1, 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step is isolating the subpopulations. 
```{r}
maledata <- filter(FinalData, sex=="guy")
maledata$sex <- as.factor(maledata$sex)
maledata$infected=factor(maledata$infected)
maledata$infected <- relevel(maledata$infected, "OK")
```

# Males

# barchart
```{r}
males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart shows that out of the approximately 340 males who were infected with covid-19, 57% of them were given the drug, and 43% of them were given the placebo. 


# Numerical summaries 
```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
The table shows that about 340 were infected with covid-19 out of the approximately 10,400 males in the study. The table shows that 3.72% of males who received the drug got covid-19, and 2.55% of the males who got the placebo contracted covid-19. The table also shows, that males got the vaccine were about 50% more likely to contract Covid-19, this suggests that the vaccine is not effective for males. 

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is so small at 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ The null hypothesis is given taking the vaccine there is no difference of whether males can contract covid-19 or not.
$H_A$ The alternative hypthoesis is that the vaccine in fact does make a difference in whether or not a male contracts covid-19. Therefore we reject the null hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 59.5% effectiveness. 

Ultimately the vaccine is 40% effective in the male population which is absolutely ineffective. 

# Females

```{r}
femaledata <- filter(FinalData, sex=="gal")
femaledata$sex <- as.factor(femaledata$sex)
femaledata$infected=factor(femaledata$infected)
femaledata$infected <- relevel(femaledata$infected, "OK")
levels(femaledata$infected)
```

# barchart

```{r}
females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart shows that out of the approximately 607 females who were infected with covid-19, 32% of them were given the drug, and 66% of them were given the placebo. 

# Numerical summaries 
```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```
The table shows that about 607 females were infected with covid-19 out of the approximately 9,200 females in the study. The table shows that 4.44% of females who received the drug got covid-19, and 8.67% of the females who got the placebo contracted covid-19. The table also shows, that females got the vaccine were about 67% more likely to contract Covid-19, this suggests that the vaccine is effective for females. 


## Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The p-value is so small at 0.000000000000002 is so small that it is highly unlikely that the results are by chance. 
$H_0$ The null hypothesis is given taking the vaccine there is no difference of whether males can contract covid-19 or not.
$H_A$ The alternative hypothesis is that the vaccine in fact does make a difference in whether or not a female contracts covid-19. Therefore we reject the null hypothesis. 

The odds ratio from the Fisher Exact test is 2.05:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 67% effectiveness. 

Ultimately the vaccine is 67% effective in the female population which is absolutely effective. 


# Drug Users

# Methods

```{r}
Druggiedata=filter(FinalData, DrugUser=="yes")
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)
Druggiedata$infected=factor(Druggiedata$infected)
Druggiedata$infected <- relevel(Druggiedata$infected, "OK")
```


# barchart

```{r}
Druggies <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart shows that out of the approximately 117 DrugUsers who were infected with covid-19, 32% of them were given the drug, and 66% of them were given the placebo. 



# Numerical summaries 
```{r}
table3 <- xtabs(~treatment + infected, data=Druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)
```
The table shows that about 117 DrugUsers were infected with covid-19 out of the approximately 2,280 DrugUsers in the study. The table shows that 33.3% of DrugUsers who received the drug got covid-19, and 66.6% of the DrugUser who got the placebo contracted covid-19. The table also shows, that DrugUsers got the vaccine were about 60% less likely to contract Covid-19, this suggests that the vaccine is effective for DrugUsers. 

## Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The p-value is so small at 0.002273 is so small that it is highly unlikely that the results are by chance. 
$H_0$ The null hypothesis is given taking the vaccine there is no difference of whether males can contract covid-19 or not.
$H_A$ The alternative hypothesis is that the vaccine in fact does make a difference in whether or not a DrugUser contracts covid-19. Therefore we reject the null hypothesis. 

The odds ratio from the Fisher Exact test is 1.83:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 67% effectiveness. 

Ultimately the vaccine is 60% effective in the DrugUser population which is absolutely ineffective.

# LGBTQ

# Methods

```{r}
LGBTQdata <- filter(FinalData, LGBTQ=="gay")
LGBTQdata$DrugUser <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected <- relevel(LGBTQdata$infected, "OK")
```

# barchart

```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart shows that out of the approximately 116 LGBTQ who were infected with covid-19, 70% of them were given the drug, and 29% of them were given the placebo. 


# Numerical summaries 
```{r}
table4 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The table shows that about 117 LGBTQ were infected with covid-19 out of the approximately 2,040 LGBTQ in the study. The table shows that 70% of LGBTQ who received the drug got covid-19, and 29% of the LGBTQ who got the placebo contracted covid-19. The table also shows, that LGBTQ got the vaccine were about 29% more likely to contract Covid-19, this suggests that the vaccine is not effective for LGBTQ population. 


## Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The p-value is so small at 0.0000057 is so small that it is highly unlikely that the results are by chance. 
$H_0$ The null hypothesis is given taking the vaccine there is no difference of whether LGBTQ can contract covid-19 or not.
$H_A$ The alternative hypothesis is that the vaccine in fact does make a difference in whether or not a LGBTQ contracts covid-19. Therefore we reject the null hypothesis. 

The odds ratio from the Fisher Exact test is 0.40:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 29% effectiveness. 

Ultimately the vaccine is 29% effective in the LGBTQ population which is absolutely ineffective.

# Discussion & Conclusion

When looking at the 4 populations, we found that the two populations that had the highest rate of effectiveness were the females at 67% and the Drugusers at 60%. Both the male and population and the LGBTQ populations showed that the vaccine was not beneficial to them and were ineffective at protecting the from contracting Covid-19, the Males had a 40% effectiveness, While the LGBTQ community had a 29% effectiveness, which is not all effective at all. 




