---
title: "Covid91 vaccine study Final2021"
author: "Thomas Matthews!"
date: "May 1, 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step is isolating the subpopulations. 
```{r}
maledata=filter(FinalData, sex=="guy")
maledata$sex <- as.factor(maledata$sex)
maledata$treatment <- as.factor(maledata$sex)
```

# Males

# bar charts
```{r}
males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart shows that out of the approximately 340 males who were infected with covid-19, 57% of them were given the drug, and 43% of them were given the placebo. 


# Numerical summaries 
```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
The table shows that about 340 were infected with covid-19 out of the approximately 10,400 males in the study. The table shows that 3.72% of males who received the drug got covid-19, and 2.55% of the males who got the placebo contracted covid-19. The table also shows, that males got the vaccine were about 50% more likely to contract Covid-19, this suggests that the vaccine is not effective for males. 

## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi squared table shows that the p-value of 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ The null hypothesis is given taking the vaccine there is no difference of whether males can contract covid-19 or not.
$H_A$ The alternative hypthoesis is that the vaccine in fact does make a difference in whether or not a male contracts covid-19. Therefore we reject the null hypothesis. 


# Females

# LGBTQ

# Drug Users

# Discussion & Conclusion






